Identification and performance of stock-recruitment functions in state

### 2024-02-07 Ecov_study/recruitment_functions/presentation/SRR_presentation.Rmd
nearly final presentation

## OM factors and simulated data examples
```{r,  echo=FALSE, message=FALSE, warnings=FALSE}

@@

@@

```{r,  echo=FALSE, message=FALSE, warnings=FALSE}

df.ems    <- readRDS(file.path(here::here(),"Ecov_study", "recruitment_functions", "inputs", "df.ems.RDS"))

@@

- ~3.3-3.5% crashes, typically for Ecov=Masking ("BH_4") and $\sigma_{R}$=0.1
```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
 knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 'crashed.models.plot_beta_fix.png') )
```

@@

## Unconverged Models (from EM having'best' EM per OM-Simulation)

```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
 knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 'bad.mods_lowestAIC.plot_grad_6_SE_100_beta_fix.png') )
```

## Unconverged Models (considering all EMs per OM-Simulation)

@@

## Unconverged Model - The Offending Parameters

@@

## Unconverged Model - Even More Offending Parameters

@@

##  $\Delta$AIC  - Median and 95% probability interval
 - Median difference in AIC (filled circle) with 2.5th and 97.5th percentiles; horizontal line at 2 for reference

@@

##  Model probability  - Median and 95% probability interval
 - Median model probability (filled circle) with 2.5th and 97.5th percentiles; horizontal line at 0.5 for reference

@@

## Assessment error - Recruitment (all years)
- Median relative error (filled circle) with 2.5th and 97.5th percentiles

@@

## Assessment error - Recruitment (last 10 years)
- Median relative error (filled circle) with 2.5th and 97.5th percentiles

@@

## Assessment error - Recruitment (terminal year)
- Median relative error (filled circle) with 2.5th and 97.5th percentiles

@@

## Assessment error - Spawning Stock Biomass (all years)
- Median relative error (filled circle) with 2.5th and 97.5th percentiles

@@

## Assessment error - Spawning Stock Biomass (terminal year)
- Median relative error (filled circle) with 2.5th and 97.5th percentiles

@@

## Assessment error - Fbar (all years)
- Median relative error (filled circle) with 2.5th and 97.5th percentiles

@@

## Assessment error - Fbar (terminal year)
- Median relative error (filled circle) with 2.5th and 97.5th percentiles

@@

## Bias of estimated parameters - Ecov $\beta$
- Median relative error (filled circle) with 2.5th and 97.5th percentiles

@@

## Bias of estimated parameters - Ecov mean
- Median relative error (filled circle) with 2.5th and 97.5th percentiles
 ```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
  knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 'RE.Ecov.mean.plot_beta_fix.png') )
  ```

## Bias of estimated parameters - $\rho_{Ecov}$
- Median relative error (filled circle) with 2.5th and 97.5th percentiles
 ```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
  knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 'RE.Ecov.rho.plot.ylim_beta_fix.png') )
  ```

## Bias of estimated parameters - $\sigma_{Ecov}$
- Median relative error (filled circle) with 2.5th and 97.5th percentiles
 ```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
  knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 'RE.Ecov_beta.plot_beta_fix.png') )
  ```

## Bias of estimated parameters - SRR a
- Median relative error (filled circle) with 2.5th and 97.5th percentiles
 ```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
  knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 'RE.mean_rec1.plot.ylim_beta_fix.png') )
  ```

## Bias of estimated parameters - SRR b
- Median relative error (filled circle) with 2.5th and 97.5th percentiles
 ```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
  knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 'RE.mean_rec2.plot.ylim_beta_fix.png') )
  ```

## Bias of estimated parameters - $\sigma_{R}$
- Median relative error (filled circle) with 2.5th and 97.5th percentiles
 ```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
  knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 'RE.R_sigma.plot.ylim_beta_fix.png') )
  ```

## Bias of estimated parameters - $\rho_{R}$
- Median relative error (filled circle) with 2.5th and 97.5th percentiles

 ```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
  knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 'RE.R_cor.plot_beta_fix.png') )
  ```

## Mohn's $\rho$

## Projections: Specifications and Ecov assumptions

## Projection performance relative to assumptions about the Ecov

@@

- EMs with only recruitment random effects generally performed similarly well as OM=EM, including for cases with an SR relationship and/or an ecov effect

@@

- This is what we suggest for follow-up
- Note that WP1-Appendix looked at $\sigma_{R}$=0.5 and found no difference from results in WP1

>>>>

## OM factors 
```{r,  echo=FALSE, message=FALSE, warnings=FALSE}

@@

## OM factor for Recruitment variability (R_sig) relative to recent assessments
```{r,  echo=FALSE, message=FALSE, warnings=FALSE}

rsig.assmt <- read.csv(file=file.path(here::here(),"Ecov_study", "recruitment_functions", "tables", "sigmaR_assessment_estimates.csv"), header=TRUE)

   knitr::kable(rsig.assmt )

@@

```{r,  echo=FALSE, message=FALSE, warnings=FALSE}
library(tidyverse)

df.ems    <- readRDS(file.path(here::here(),"Ecov_study", "recruitment_functions", "inputs", "df.ems.RDS"))
df.ems2 <- cbind(EM=seq(1,6), df.ems)

@@

- ~3.3-3.5% crashes, typically for Ecov=Masking ("BH_4") and $\sigma_{R}$=0.1
- most problematic for Masking (BH_4) and exacerbated by $\sigma_{R}$=0.1
```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
 knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 'crashed.models.plot_beta_fix.png') )
```

@@

## Unconverged Models (from EM having'best' EM per OM-Simulation)
- most problematic for $\sigma_{R}$=0.1; BH with no Ecov and high fishing contrast (H-MSY) and mean SRR with no contrast (MSY)
```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
 knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 'bad.mods_lowestAIC.plot_grad_6_SE_100_beta_fix.png') )
```

## Unconverged Models (considering all EMs per OM-Simulation)
- most problematic for Beverton-Holt SRR, worse with no contrast in fishing history (MSY)

@@

## Unconverged Model - The Offending Parameters 
- the SRR recruitment parameters led to most convergence failures

@@

## Unconverged Model - Second Most Offending Parameters

- the Ecov parameters led to the second most convergence failures

@@

## Results layout (example)
- Plots show median relative error (filled circle) with 2.5th and 97.5th percentiles
- OM factors as facets (top and right side)
- EM models on the x-axis
- Blue indicates EM=OM (diagonal of first 4 columns w/in facet blocks), incorrect EMs in red

 ```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='65%', fig.align='center'}
  knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 're.rec.all.yrs.plot_beta_fix.png') )
  ```

##  $\Delta$AIC  - Median and 95% probability interval
- Smaller $\Delta$AIC when $\sigma_{R}$ is larger

@@

##  Model probability  - Median and 95% probability interval
- True OMs receive more probability with $\sigma_{R}$=0.1 and high fishing contrast; otherwise, mean SRR received more probability

@@

## Assessment error - Recruitment (all years)
- Median unbiased, $\sigma_{R}$ determined quantile range

@@

## Assessment error - Recruitment (last 10 years)
- Median unbiased, $\sigma_{R}$ determined quantile range

@@

## Assessment error - Recruitment (terminal year)
- Median unbiased, $\sigma_{R}$ determined quantile range

@@

## Assessment error - Spawning Stock Biomass (all years)
- Median unbiased, $\sigma_{R}$ determined quantile range

@@

## Assessment error - Spawning Stock Biomass (terminal year)
- Median unbiased, $\sigma_{R}$ determined quantile range

@@

## Assessment error - Fbar (all years)
- Median unbiased, $\sigma_{R}$ determined quantile range

@@

## Assessment error - Fbar (terminal year)
- Median unbiased, $\sigma_{R}$ determined quantile range

@@

## Bias of estimated parameters - Ecov $\beta$

- Limiting and Masking had worst performance (BH_3 and BH_4)

@@

## Bias of estimated parameters - Ecov mean
- Larger quantile range when Ecov correlation is high (0.8)
 ```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
  knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 'RE.Ecov.mean.plot_beta_fix.png') )
  ```

## Bias of estimated parameters - $\rho_{Ecov}$
- Well estimated when Ecov correlation was high (0.8)
 ```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
  knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 'RE.Ecov.rho.plot.ylim_beta_fix.png') )
  ```

## Bias of estimated parameters - $\sigma_{Ecov}$
- Positively biased when Ecov correlation is high (0.8)
 ```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
  knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 'RE.Ecov.sigma.plot_beta_fix.png') )
  ```

## Bias of estimated parameters - SRR a
- Unbiased when $\sigma_{R}$ =0.1 and high fishing contrast; otherwise biased +/- 0.2 (BH_4 could be much worse)
 ```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
  knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 'RE.mean_rec1.plot.ylim_beta_fix.png') )
  ```

## Bias of estimated parameters - SRR b
- Unbiased when $\sigma_{R}$ =0.1 and high fishing contrast; otherwise biased +/- 0.2 (BH_4 could be much worse)

 ```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
  knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 'RE.mean_rec2.plot.ylim_beta_fix.png') )
  ```

## Bias of estimated parameters - $\sigma_{R}$
- Bias pattern mirroring between $\sigma_{R}$ and correlation
 ```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
  knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 'RE.R_sigma.plot.ylim_beta_fix.png') )
  ```

## Bias of estimated parameters - $\rho_{R}$
- Bias pattern mirroring between $\sigma_{R}$ and correlation
- For most informative case (H-MSY, $\sigma_{R}$=0.1, low recruitment correlation), mean SRR has large positive bias
 ```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
  knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 'RE.R_cor.plot_beta_fix.png') )
  ```

## Mohn's $\rho$ - Recruitment (7 year peel)  
- Missing symbol indicates failed peels for one or more EM fits
- $\sigma_{R}$ and observation error were influential factors
 ```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
  knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 'recr.rho_beta_fix.png') )
  ```

## Mohn's $\rho$ - SSB  (7 year peel)   
- observation error was influential factor

 ```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
  knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 'ssb.rho_beta_fix.png') )
  ```

## Mohn's $\rho$ - Fbar  (7 year peel) 
- observation error was influential factor

 ```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
  knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 'fbar.rho_beta_fix.png') )
  ```

## Mohn's $\rho$ - Classification Tree for Recruitment, SSB, Fbar
-

 ```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
  knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 'reg_tree_Rho_beta_fix.pdf') )
  ```

## Mohn's $\rho$ - Random Effects on Recruitment (7 year peel)
- Negative bias when $\sigma_{R}$ is low
- Expected worse $\rho$ from incorrect EM, but was not the case

 ```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
  knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 'recr.dev.rho_beta_fix.png') )
  ```

## Mohn's $\rho$ - Classification Tree for Random Effects on Recruitment

 ```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
  knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 'reg_tree_Rdev_retro_beta_fix.pdf') )
  ```

## Projections: Specifications and Ecov assumptions 
- Projections were made from the 10th peel of each fitted model (Terminal year - 10)
- A 5-year average of maturity, weight at age, and selectivity specified (time-invariant anyway)
- F in the projection was set at the true OM F value
- The environmental index was treated 3 ways

    \begin{enumerate}
      \item A 5-year average of the most recent index value was held constant in the 10 year projection ('Avg5yr')
      \item The estimated process continued in the 10 year projection ('Continue')
      \item The observed environmental index was specified in the 10 year projection ('ObsEcov') 
    \end{enumerate}

- Projected Recruitment, SSB, and Catch were compared to the true OM values in the projection

## Projection performance - Relative Error Recruitment (first 3 years)
- Minimal bias (median at most -0.1 for mean SRR)
- Quantile range much larger for BH
 ```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
  knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 're.rec.proj.plot_beta_fix.png') )
    ```

## Projection performance - Relative Error Recruitment (all 10 years)
- Minimal bias (median at most -0.1 for mean SRR)
- Quantile range much larger for BH

 ```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
  knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 're.rec.proj.plot_beta_fix.png') )
  ```

## Projection performance - Recruitment Uncertainty (first 3 years)
- CV slightly lower for mean SRR, usually narrower quantile range

 ```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
  knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 'sd.rec.proj.plot3yr_beta_fix.png') )
    ```

## Projection performance - Recruitment Uncertainty (all 10 years)
- CV slightly lower for mean SRR, usually narrower quantile range

 ```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
  knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 'sd.rec.proj.plot_beta_fix.png') )
  ```

## Projection performance - Relative Error SSB (first 3 years)
- Minimal bias (median at most -0.1 for mean SRR)
- Quantile range much larger for BH

 ```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
  knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 're.ssb.proj.plot_beta_fix.png') )
    ```

## Projection performance - Relative Error SSB (all 10 years)
- Minimal bias (median at most -0.1 for mean SRR)
- Quantile range much larger for BH

 ```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
  knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 're.ssb.proj.plot_beta_fix.png') )
  ```

## Projection performance - SSB Uncertainty (first 3 years)
- CV slightly lower for mean SRR, usually narrower quantile range

 ```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
  knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 'sd.ssb.proj.plot3yr_beta_fix.png') )
    ```

## Projection performance - SSB Uncertainty (all 10 years)
- CV slightly lower for mean SRR, usually narrower quantile range

 ```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
  knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 'sd.ssb.proj.plot_beta_fix.png') )
  ```

## Projection performance - Relative Error Catch (first 3 years)
- Minimal bias (median at most -0.1 for mean SRR)
- Quantile range much larger for BH

 ```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
  knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 're.catch.proj.plot_beta_fix.png') )
    ```

## Projection performance - Relative Error Catch (all 10 years)
- Minimal bias (median at most -0.1 for mean SRR)
- Quantile range much larger for BH

 ```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
  knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 're.catch.proj.plot_beta_fix.png') )
  ```

## Projection performance - Catch Uncertainty (first 3 years)
- CV slightly lower for mean SRR, usually narrower quantile range

 ```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
  knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 'sd.catch.proj.plot3yr_beta_fix.png') )
    ```

## Projection performance - Catch Uncertainty (all 10 years)
- CV slightly lower for mean SRR, usually narrower quantile range

 ```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
  knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 'sd.catch.proj.plot_beta_fix.png') )
  ```

## Projection Summary
- Minimal difference in median bias
- Recruitment is highly uncertain in projections, exacerbated by length of projection
- SSB and Catch also show increased uncertainty in longer projections
- Long projections are expected to be very imprecise, recommend shorter projections
- Tighter quantile range and slightly lower CV for mean SRR

@@

- EMs with mean SRR and recruitment random effects generally performed similarly well as OM=EM, including for cases with an SR relationship and/or an ecov effect

@@

- In WP-1, we recommended exploring additional levels of $\sigma_{R}$=0.5; however, we note that WP1-Appendix looked at $\sigma_{R}$=0.5 and found no difference
